---
title: "Data merging and QC script"
author: c("Oliver J. Lechtenfeld", "James Stegen")
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(here)
library(tidyverse)
library(ggpubr)
library(ggridges)
library(googledrive)

```


```{r authenticate google}

drive_auth() # use the same Google Account used to access the Google Drive Folder


```


# Read Input Files from Google Drive

Please update the links, if a different file shall be used!
```{r download csv files into working directory}

# original data
drive_download(as_id("https://drive.google.com/file/d/1BglQCdkUHMyP0vCLWvVzUSJ2K5m-LCaV/view?usp=drive_link"), overwrite = T) # Merged_Processed_Data - 10-19-23.csv
drive_download(as_id("https://drive.google.com/file/d/1MtbAViGO9Cw8B9p0mx6gcjda-DgXJYr9/view?usp=drive_link"), overwrite = T) # Merged_Processed_Mol - 10-19-23.csv

# trimmed data
drive_download(as_id("https://drive.google.com/file/d/1AaJYEK0IM5TiZL2YeCmKOIdCaIPipEYi/view?usp=drive_link"), overwrite = T) # DOM_Synthesis_Data_Trim.csv
drive_download(as_id("https://drive.google.com/file/d/135aP9tmAcXS33nE9atYztIXZ1U3DknvE/view?usp=drive_link"), overwrite = T) # DOM_Synthesis_Mol_Trim.csv
drive_download(as_id("https://drive.google.com/file/d/1Gb-Z2IwHD127XVDHuXiYZZgmky2nBR3C/view?usp=drive_link"), overwrite = T) # DOM_Synthesis_Mol_Trim_Lambda.csv

# aggregated data
drive_download(as_id("https://drive.google.com/file/d/1KSM47o6vfB4c1sjG0kLkoVMc3GhLZxHT/view?usp=drive_link"), overwrite = T) # Summary_Andrew_crosstab_wihout_iso_na.csv
drive_download(as_id("https://drive.google.com/file/d/1fVbjuFVfQTnYgkaBstWlwWzs1MKT-4Rm/view?usp=drive_link"), overwrite = T) # Summary_Andrew_crosstab_wihout_iso_wa.csv

# metadata
drive_download(as_id("https://docs.google.com/spreadsheets/d/10PUYmIsTrmFiBefTAeog6ovh_aHaxGcn/edit?usp=drive_link"), overwrite = T) # AH_metadata_merge.xlsx



```



# Read original data
Using Bobs code
```{r read untrimed data, cache=T}

# data = read.csv("Merged_Processed_Data - 10-19-23.csv", check.names = F, row.names = 1)
data = data.frame(fread("Merged_Processed_Data - 10-19-23.csv", check.names = F), row.names = 1)
# mol = read.csv("Merged_Processed_Mol - 10-19-23.csv", check.names = F, row.names = 1)
mol = data.frame(fread("Merged_Processed_Mol - 10-19-23.csv", check.names = F), row.names = 1)

# Breaking apart by dataset
factors = data.frame(Sample = colnames(data),
                     Dataset = str_extract(colnames(data), "[^_]+"))

```



```{r prepare data for metric distribution, cache=T}
# Create empty object
dataset.data = data.frame(matrix(nrow = nrow(data), ncol = length(unique(factors$Dataset)),
                                 dimnames = list(row.names(data),
                                                 unique(factors$Dataset))),
                          check.names = F)

# Summarize presence/absence by dataset
for(i in 1:ncol(dataset.data)){
  # current dataset
  curr.data = colnames(dataset.data)[i]
  
  # Select matching samples
  temp.data = data[,which(factors$Dataset %in% curr.data)]
  
  # add in sums to pregenerated object
  dataset.data[,i] = rowSums(temp.data)
  
  # cleanup
  rm(temp.data, curr.data)
}

# Convert to presence/absence
dataset.data[dataset.data > 0] = 1

# Generate mol file with dataset information
dataset.mol = NULL

for(i in 1:ncol(dataset.data)){
  # find formulas
  w = which(dataset.data[,i] > 0)
  
  # find corresponding molecular information
  temp = mol[w,]
  
  # add dataset information
  temp$Dataset = colnames(dataset.data)[i]
  
  # add in to dataset.mol
  dataset.mol = rbind(dataset.mol, temp)
  
}
```


```{r plotting metric distribution, message=FALSE, warning=FALSE}

# plotting metric distribution
dataset.mol %>% select(Dataset, AI_Mod, DBE_1, NOSC, OtoC_ratio, HtoC_ratio) %>%
  gather(Variable, Value, -Dataset) %>%
  ggplot(aes(x = Value, y = Dataset))+
  geom_density_ridges(aes(fill = Dataset), alpha = 0.5)+
  facet_wrap(.~Variable, scales = "free")+
  xlab("Derived DOM Metric")+
  theme_bw()


```
# Trim data

Using James's script; only run, in case smth has to be changed
```{r prepare trimed data, eval=FALSE, include=FALSE}

# QC data matrix

formula.drop.matrix = data.frame('singleton' = rep(NA,nrow(mol)),'P'= rep(NA,nrow(mol)), 'HtoC' = rep(NA,nrow(mol)), 'OtoC' = rep(NA,nrow(mol)), 'DBE_1' = rep(NA,nrow(mol)), 'DBE_O' = rep(NA,nrow(mol)), 'mass' = rep(NA,nrow(mol)),row.names = rownames(mol))

# define singletons
formula.occurence = rowSums(x = data)
head(formula.occurence)
singletons = formula.occurence[which(formula.occurence <= 1)]
head(singletons)
formula.drop.matrix$singleton[which(rownames(formula.drop.matrix) %in% names(singletons))] = 1

# define P formulas
p.formulas = rownames(mol)[which(mol$P > 0)]
formula.drop.matrix$P[which(rownames(formula.drop.matrix) %in% p.formulas)] = 1

# define masses out of range
bad.mass.formulas = rownames(mol)[which(as.numeric(rownames(mol)) < 200 | as.numeric(rownames(mol)) > 800)]
formula.drop.matrix$mass[which(rownames(formula.drop.matrix) %in% bad.mass.formulas)] = 1

# define O:C 
OC.formulas = rownames(mol)[which(mol$OtoC_ratio > 1)]
formula.drop.matrix$OtoC[which(rownames(formula.drop.matrix) %in% OC.formulas)] = 1

# define H:C
HC.formulas = rownames(mol)[which(mol$HtoC_ratio > 2.5 | mol$HtoC_ratio < 0.3)]
formula.drop.matrix$HtoC[which(rownames(formula.drop.matrix) %in% HC.formulas)] = 1

# define DBE minus C
DBE.C.formulas = rownames(mol)[which(mol$DBE_1 <= I(mol$C*0.6 - 15))]
formula.drop.matrix$DBE_1[which(rownames(formula.drop.matrix) %in% DBE.C.formulas)] = 1

# define DBE minus O
DBE.O.formulas = rownames(mol)[which(mol$DBE_O < I(-10) | mol$DBE_O > 10)]
formula.drop.matrix$DBE_O[which(rownames(formula.drop.matrix) %in% DBE.O.formulas)] = 1

# number of drops per filter
colSums(x = formula.drop.matrix,na.rm = T)
#singleton         P      HtoC      OtoC     DBE_1     DBE_O      mass 
#80211    104223       244      4941     20441     92566     54028 

# number of filters per formula
filters.per.formula = rowSums(x = formula.drop.matrix,na.rm = T)
range(filters.per.formula)
formula.to.drop = names(filters.per.formula)[which(filters.per.formula > 0)]
length(formula.to.drop)

# trim down mol
mol.trim = mol[-which(rownames(mol) %in% formula.to.drop),]
dim(mol.trim)
summary(mol.trim)

# make Van-K plot
plot(mol.trim$HtoC_ratio ~ mol.trim$OtoC_ratio,cex=0.3)

# trim down the data file
data.trim = data[which(rownames(data) %in% rownames(mol.trim)),]
dim(data.trim)

# look at formula richness distribution
sample.richness = colSums(x = data.trim)
hist(sample.richness)

# needs to be True
identical(x = row.names(data.trim), y = row.names(mol.trim))

# write out files
# fwrite(x = mol.trim, file = "DOM_Synthesis_Mol_Trim.csv")
# fwrite(x = data.trim, file = "DOM_Synthesis_Data_Trim.csv")

rm(formula.drop.matrix)
rm(data, mol,dataset.data, dataset.mol)
rm(bad.mass.formulas, DBE.C.formulas, DBE.O.formulas, filters.per.formula, formula.occurence, formula.to.drop, HC.formulas, OC.formulas, p.formulas, singletons, w)

```


# Read final, trimmed data
using James code

```{r read trimed data, cache=T}

# data = read.csv("Merged_Processed_Data - 10-19-23.csv", check.names = F, row.names = 1)
data.trim = data.frame(fread("DOM_Synthesis_Data_Trim.csv", check.names = F), row.names = 1)
# mol = read.csv("Merged_Processed_Mol - 10-19-23.csv", check.names = F, row.names = 1)
mol.trim = data.frame(fread("DOM_Synthesis_Mol_Trim.csv", check.names = F), row.names = 1)

# Breaking apart by dataset
factors.trim = data.frame(Sample = colnames(data.trim),
                     Dataset = str_extract(colnames(data.trim), "[^_]+"))

```


```{r prepare trimed data for metric distribution, cache=T}
# Create empty object
dataset.data.trim = data.frame(matrix(nrow = nrow(data.trim), ncol = length(unique(factors.trim$Dataset)),
                                 dimnames = list(row.names(data.trim),
                                                 unique(factors.trim$Dataset))),
                          check.names = F)

# Summarize presence/absence by dataset
for(i in 1:ncol(dataset.data.trim)){
  # current dataset
  curr.data = colnames(dataset.data.trim)[i]
  
  # Select matching samples
  temp.data = data.trim[,which(factors.trim$Dataset %in% curr.data)]
  
  # add in sums to pregenerated object
  dataset.data.trim[,i] = rowSums(temp.data)
  
  # cleanup
  rm(temp.data, curr.data)
}

# Convert to presence/absence
dataset.data.trim[dataset.data.trim > 0] = 1

# Generate mol file with dataset information
dataset.mol.trim = NULL

for(i in 1:ncol(dataset.data.trim)){
  # find formulas
  w = which(dataset.data.trim[,i] > 0)
  
  # find corresponding molecular information
  temp = mol.trim[w,]
  
  # add dataset information
  temp$Dataset = colnames(dataset.data.trim)[i]
  
  # add in to dataset.mol
  dataset.mol.trim = rbind(dataset.mol.trim, temp)
  
}
```

```{r plotting metric distribution, message=FALSE, warning=FALSE}

# plotting metric distribution
dataset.mol.trim %>% select(Dataset, AI_Mod, DBE_1, NOSC, OtoC_ratio, HtoC_ratio) %>%
  gather(Variable, Value, -Dataset) %>%
  ggplot(aes(x = Value, y = Dataset))+
  geom_density_ridges(aes(fill = Dataset), alpha = 0.5)+
  facet_wrap(.~Variable, scales = "free")+
  xlab("Derived DOM Metric")+
  theme_bw()


```



# Merge Data and Mol Tables
```{r merge data and mol}


# quick check
sum(is.na(data.trim))

tmp <- as.data.table(data.trim)[, mz := row.names(data.trim)]

tmp2 <- as.data.table(mol.trim)[, mz := row.names(mol.trim)]

tmp3 <- merge(tmp2, tmp, by = "mz")

# Replace zeros with "NA"
cols_to_replace <- colnames(data.trim)

tmp3[, (cols_to_replace) := lapply(.SD, function(x) ifelse(x == 0, NA, x)), .SDcols = cols_to_replace]

# renormalize to 0:1000

tmp3[, (cols_to_replace) := lapply(.SD, function(x) scales::rescale(x, to = c(0, 1000))), .SDcols = cols_to_replace]

tmp4 <- melt(tmp3, id.vars = colnames(tmp3)[!colnames(tmp3) %in% cols_to_replace], value.name = "Intensity", variable.name = "Sample", measure.vars = cols_to_replace, na.rm = T)
# tmp <- tmp[intensity == 0,]

tmp4[, Dataset := str_extract(Sample, "[^_]+")]

tmp4[is.infinite(Intensity), .N]

# merged data table
data.mol.trim <- copy(tmp4)


```

Note that some samples had missing MF at this stage of data processing and were removed from the data

# MF Distribution
```{r}

ggplot(data.mol.trim[, .N, by = c("Dataset", "Sample")], aes(x = Dataset, y = N, fill = Dataset)) + geom_violin() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data.mol.trim[, .N, by = c("Sample")], aes(y = N, x = "All Data")) + geom_violin(trim = T) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

Molecular_number_threshold <- 500

```



# Heteroatom Class Distributions
```{r factors for sorting}

fctr_lvls_cho <- c("O1",
"O2",
"O3",
"O4",
"O5",
"O6",
"O7",
"O8",
"O9",
"O10",
"O11",
"O12",
"O13",
"O14",
"O15",
"O16",
"O17",
"O18",
"O19",
"O20",
"O21",
"O22",
"O23",
"O24",
"O25")

fctr_lvls_n <- c("N1 O1",
"N1 O2",
"N1 O3",
"N1 O4",
"N1 O5",
"N1 O6",
"N1 O7",
"N1 O8",
"N1 O9",
"N1 O10",
"N1 O11",
"N1 O12",
"N1 O13",
"N1 O14",
"N1 O15",
"N1 O16",
"N1 O17",
"N1 O18",
"N1 O19",
"N1 O20",
"N1 O21",
"N1 O22",
"N1 O23",
"N1 O24",
"N1 O25",
"N2 O1",
"N2 O2",
"N2 O3",
"N2 O4",
"N2 O5",
"N2 O6",
"N2 O7",
"N2 O8",
"N2 O9",
"N2 O10",
"N2 O11",
"N2 O12",
"N2 O13",
"N2 O14",
"N2 O15",
"N2 O16",
"N2 O17",
"N2 O18",
"N2 O19",
"N2 O20",
"N2 O21",
"N2 O22",
"N2 O23",
"N2 O24",
"N2 O25")

fctr_lvls_s <- c("S1 O1",
"S1 O2",
"S1 O3",
"S1 O4",
"S1 O5",
"S1 O6",
"S1 O7",
"S1 O8",
"S1 O9",
"S1 O10",
"S1 O11",
"S1 O12",
"S1 O13",
"S1 O14",
"S1 O15",
"S1 O16",
"S1 O17",
"S1 O18",
"S1 O19",
"S1 O20",
"S1 O21",
"S1 O22",
"S1 O23",
"S1 O24",
"S1 O25",
"S2 O1",
"S2 O2",
"S2 O3",
"S2 O4",
"S2 O5",
"S2 O6",
"S2 O7",
"S2 O8",
"S2 O9",
"S2 O10",
"S2 O11",
"S2 O12",
"S2 O13",
"S2 O14",
"S2 O15",
"S2 O16",
"S2 O17",
"S2 O18",
"S2 O19",
"S2 O20",
"S2 O21",
"S2 O22",
"S2 O23",
"S2 O24",
"S2 O25")

fctr_lvls_ns <- c("N1 S1 O1",
"N1 S1 O2",
"N1 S1 O3",
"N1 S1 O4",
"N1 S1 O5",
"N1 S1 O6",
"N1 S1 O7",
"N1 S1 O8",
"N1 S1 O9",
"N1 S1 O10",
"N1 S1 O11",
"N1 S1 O12",
"N1 S1 O13",
"N1 S1 O14",
"N1 S1 O15",
"N1 S1 O16",
"N1 S1 O17",
"N1 S1 O18",
"N1 S1 O19",
"N1 S1 O20",
"N1 S1 O21",
"N1 S1 O22",
"N1 S1 O23",
"N1 S1 O24",
"N1 S1 O25",
"N2 S1 O1",
"N2 S1 O2",
"N2 S1 O3",
"N2 S1 O4",
"N2 S1 O5",
"N2 S1 O6",
"N2 S1 O7",
"N2 S1 O8",
"N2 S1 O9",
"N2 S1 O10",
"N2 S1 O11",
"N2 S1 O12",
"N2 S1 O13",
"N2 S1 O14",
"N2 S1 O15",
"N2 S1 O16",
"N2 S1 O17",
"N2 S1 O18",
"N2 S1 O19",
"N2 S1 O20",
"N2 S1 O21",
"N2 S1 O22",
"N2 S1 O23",
"N2 S1 O24",
"N2 S1 O25",
"N1 S2 O1",
"N1 S2 O2",
"N1 S2 O3",
"N1 S2 O4",
"N1 S2 O5",
"N1 S2 O6",
"N1 S2 O7",
"N1 S2 O8",
"N1 S2 O9",
"N1 S2 O10",
"N1 S2 O11",
"N1 S2 O12",
"N1 S2 O13",
"N1 S2 O14",
"N1 S2 O15",
"N1 S2 O16",
"N1 S2 O17",
"N1 S2 O18",
"N1 S2 O19",
"N1 S2 O20",
"N1 S2 O21",
"N1 S2 O22",
"N1 S2 O23",
"N1 S2 O24",
"N1 S2 O25",
"N2 S2 O1",
"N2 S2 O2",
"N2 S2 O3",
"N2 S2 O4",
"N2 S2 O5",
"N2 S2 O6",
"N2 S2 O7",
"N2 S2 O8",
"N2 S2 O9",
"N2 S2 O10",
"N2 S2 O11",
"N2 S2 O12",
"N2 S2 O13",
"N2 S2 O14",
"N2 S2 O15",
"N2 S2 O16",
"N2 S2 O17",
"N2 S2 O18",
"N2 S2 O19",
"N2 S2 O20",
"N2 S2 O21",
"N2 S2 O22",
"N2 S2 O23",
"N2 S2 O24",
"N2 S2 O25")

```

```{r plot distribution by major element, message=FALSE, warning=FALSE}

het_check <- data.mol.trim[, .N, by = c("Heteroatom.Class", "Sample", "Dataset")][, .(mean_count = mean(N)), by = c("Heteroatom.Class", "Dataset")]

test <- copy(het_check)

for (het in c("CHO", "N", "S", "N+S")) {

  if (het == "CHO") {
  
      test_data <- test[!grepl(pattern = "N", Heteroatom.Class) & !grepl(pattern = "S", Heteroatom.Class), ] %>% dcast(., Heteroatom.Class~Dataset, value.var = "mean_count" , fill = 0)
            test_data[, Heteroatom.Class := factor(Heteroatom.Class, levels = fctr_lvls_cho)]

  }
  
    if (het == "N") {
  
      test_data <- test[grepl(pattern = "N", Heteroatom.Class) & !grepl(pattern = "S", Heteroatom.Class), ] %>% dcast(., Heteroatom.Class~Dataset, value.var = "mean_count" , fill = 0)
            test_data[, Heteroatom.Class := factor(Heteroatom.Class, levels = fctr_lvls_n)]

    }
  
      if (het == "S") {
  
      test_data <- test[!grepl(pattern = "N", Heteroatom.Class) & grepl(pattern = "S", Heteroatom.Class), ] %>% dcast(., Heteroatom.Class~Dataset, value.var = "mean_count" , fill = 0)
            test_data[, Heteroatom.Class := factor(Heteroatom.Class, levels = fctr_lvls_s)]

      }
  
      if (het == "N+S") {
  
      test_data <- test[grepl(pattern = "N", Heteroatom.Class) & grepl(pattern = "S", Heteroatom.Class), ] %>% dcast(., Heteroatom.Class~Dataset, value.var = "mean_count" , fill = 0)
      test_data[, Heteroatom.Class := factor(Heteroatom.Class, levels = fctr_lvls_ns)]

}

setkeyv(test_data, cols = "Heteroatom.Class")

print(heatmap(as.matrix(test_data, rownames = "Heteroatom.Class"), Rowv = NA, Colv = NA, revC = T, ylab = "Heteroatom.Class", xlab = "Dataset", scale = "none", main = het, keep.dendro = F))

  }

```

# Read Aggregated Values
using Dings code

```{r read aggregated values}

# load files and do some renaming + remove samples with 0 MF
# Weighted averages
data.mol.trim.agg_wa <- fread("Summary_Andrew_crosstab_wihout_iso_wa.csv")
names(data.mol.trim.agg_wa)[which(names(data.mol.trim.agg_wa ) == "variable")] <- "Sample"
data.mol.trim.agg_wa <- data.mol.trim.agg_wa[Molecular_number > 0] %>% .[, Dataset := str_extract(Sample, "[^_]+")]

# Number averages
data.mol.trim.agg_na <- fread("Summary_Andrew_crosstab_wihout_iso_na.csv")
names(data.mol.trim.agg_na)[which(names(data.mol.trim.agg_na) == "variable")] <- "Sample"
data.mol.trim.agg_na <- data.mol.trim.agg_na[Molecular_number > 0] %>% .[, Dataset := str_extract(Sample, "[^_]+")]
names(data.mol.trim.agg_na) <- sub("_wa", "_na", names(data.mol.trim.agg_na))


```

# Read Metadata
```{r read metadata}

metadata <- fread("AH_metadata_merge.csv", na.strings = c("NA", ""))

# filter data without coreMS filename

metadata <- metadata[!is.na(coreMSname)]

# renaming

metadata[, Sample := coreMSname]


```


```{r merge metadata with aggregated data}

data.mol.trim.agg_na[, Sample := sub(".csv", "", Sample)]
data.mol.trim.agg_wa[, Sample := sub(".csv", "", Sample)]

data.mol.trim.agg_na.meta <- merge(data.mol.trim.agg_na, metadata, by = "Sample")
data.mol.trim.agg_wa.meta <- merge(data.mol.trim.agg_wa, metadata, by = "Sample")

```

The number of matches is lower than in the original files!

```{r overview}

data.mol.trim.agg_na.meta[, .N, by = "type"]


```


# QC plots
## Weighted and number averaged per Dataset
```{r plot wa distributions per dataset, message=FALSE, warning=FALSE}

test <- melt(data.mol.trim.agg_wa, value.name = "wa_value", id.vars = c("Dataset", "Sample"))

ggplot(test, aes(x = Dataset, y = wa_value)) + geom_boxplot() + facet_wrap("variable", scales = "free") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
```

```{r plot na distributions per dataset, message=FALSE, warning=FALSE}

test <- melt(data.mol.trim.agg_na, value.name = "wa_value", id.vars = c("Dataset", "Sample"))

ggplot(test, aes(x = Dataset, y = wa_value)) + geom_boxplot() + facet_wrap("variable", scales = "free") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
```

```{r correlate na with wa}

cols_to_use <- c("H.C", "O.C", "MZ", "AI.mod")

for (n in cols_to_use) {
  
  test <- merge(data.mol.trim.agg_na[Molecular_number > Molecular_number_threshold, .(Sample, Dataset, .SD), .SDcols = paste0(n, "_na")], data.mol.trim.agg_wa[Molecular_number > Molecular_number_threshold, .(Sample, .SD), .SDcols = paste0(n, "_wa")], by = "Sample")
  names(test) <- sub(".SD.", "", names(test))
  
  print(ggplot(test, aes(x = get(paste0(n, "_na")), y = get(paste0(n, "_wa")), color = Dataset)) + geom_point() + labs(title = n))
  
  
}



```



## Main Molecular Descriptors
```{r overall sample distributions}

# check dependency on # Assignments
ggplot(data.mol.trim.agg_wa, aes(x = Molecular_number, y = MZ_wa, color = Dataset)) + geom_point() + labs(title = paste0("All Samples"))
ggplot(data.mol.trim.agg_wa[Molecular_number >= Molecular_number_threshold], aes(x = O.C_wa, y = H.C_wa, color = Dataset)) + geom_point() + labs(title = paste0("Samples with more than ", Molecular_number_threshold, " assignments"))
ggplot(data.mol.trim.agg_wa[Molecular_number < Molecular_number_threshold], aes(x = O.C_wa, y = H.C_wa, color = Dataset)) + geom_point() + labs(title = paste0("Samples with less than ", Molecular_number_threshold, " assignments"))

```




```{r}
library(LambdaMineR)
set_LM_plot_style()

formulas.trim <- copy(data.mol.trim)
names(formulas.trim)[which(names(formulas.trim) == "Sample")] <- "measurement_name" 
names(formulas.trim)[which(names(formulas.trim) == "OtoC_ratio")] <- "formula_oc"
names(formulas.trim)[which(names(formulas.trim) == "HtoC_ratio")] <- "formula_hc" 
names(formulas.trim)[which(names(formulas.trim) == "Intensity")] <- "peak_intensity" 
names(formulas.trim)[which(names(formulas.trim) == "Molecular.Formula")] <- "formula_string" 

plot_vK(formulas.trim, "ManyFiles_Nelha_054_01_23226.corems", color_var = "peak_intensity")

```

